---
title: "Chocolate Chip Cookies"
execute:
  error: true
author: "Joe Stuedemann"
output: html_document
---

## Reading In the Data

First, read in the CSV data of cookie ingredients.
Make sure that your end-result data has appropriate types for each column - these should match the types provided in the documentation in the README.md file.

```{r}
library(tidyverse)

choc_chip_cookie_ingredients <- read_csv("choc_chip_cookie_ingredients.csv")

choc_chip_cookie_ingredients <- choc_chip_cookie_ingredients %>%
  mutate(
    Ingredient = as.character(Ingredient),
    Text = as.character(Text),
    Recipe_Index = as.character(Recipe_Index),
    Rating = as.numeric(Rating),
    Quantity = as.numeric(Quantity),
    Unit = as.character(Unit)
  )

glimpse(choc_chip_cookie_ingredients)

```

```{python}
import pandas as pd

# 1. Load data
cookies = pd.read_csv('choc_chip_cookie_ingredients.csv')

# 2. Type Conversions
cookies['Ingredient'] = cookies['Ingredient'].astype(str)
cookies['Text'] = cookies['Text'].astype(str)
cookies['Recipe_Index'] = cookies['Recipe_Index'].astype(str)
cookies['Rating'] = pd.to_numeric(cookies['Rating'], errors='coerce')
cookies['Quantity'] = pd.to_numeric(cookies['Quantity'], errors='coerce')
cookies['Unit'] = cookies['Unit'].astype(str) 

# 3. Me verifying that this worked
print(cookies.dtypes)

```


## Exploratory Data Analysis

Exploratory data analysis is the process of getting familiar with your dataset. To get started, [this blog post](https://www.mrdbourke.com/a-gentle-introduction-to-exploratory-data-analysis/) provides a nice checklist to get you thinking:

> 1.  What question(s) are you trying to solve (or prove wrong)?
> 2.  What kind of data do you have and how do you treat different types?
> 3.  What's missing from the data and how do you deal with it?
> 4.  Where are the outliers and why should you care about them?
> 5.  How can you add, change or remove features to get more out of your data?

### Generating Questions

Generate at least 5 questions you might explore using this database of cookie ingredients.

1. What are the "essential ingredients" or the ones that appear in say 85% of all the recipes and therefore what are the rarer ones as well? 
2. Are there any outlier recipes with unusual quantities? 
3. What is the salt to sugar ratio in high rated recipes compared to low rated recipes? 
4. Is there any correlation between the amount of chocolate chips and the overall recipe rating? 
5. How do the ingredient profiles differ between the sources? (Epicurious vs Allrecipes) 


### Skimming the Data

One thing we often want to do during EDA is to examine the quality of the data - are there missing values? What quirks might exist in the dataset?

The `skimr` package in R, and the similar `skimpy` package in python (which has a much better name, in my opinion), can help provide visual summaries of the data. 

Install both packages, and read the package documentation ([R](https://cran.r-project.org/web/packages/skimr/vignettes/skimr.html), [Python](https://pypi.org/project/skimpy/)).

[Part 1] Use each package and generate summaries of your data that require the use of at least some non-default options in each package's `skim` function.


```{r}
library(skimr)

# I'm selecting only specific columns to keep it clean
choc_chip_cookie_ingredients %>%
  skim(Rating, Quantity) %>%
  yank("numeric") # I looked up the 'yanks' command to give me the numeric table specifically so it displays fully
```

```{python}
from skimpy import skim

# Non-default approach: 
skim(cookies[['Ingredient', 'Rating', 'Quantity']].iloc[:500])

# Not sure if this is right in terms of skimming but I think I'm able to skim through so I guess that's what matters.  

```

[Part 2] Write 1-2 sentences about what you can tell from each summary display you generate. Did you discover anything new about the data?

From the R summary, I see that there are 1,010 missing ratings which is kind of big deal when it comes to comparing recipes based off of what ingredients they use. I don't you'd be able to truly narrow down "What makes a good cookie?". The other thing I found in the python summary is the standard deviation for quantity is larger than the mean (2.53 compared to 2.23). This just means that I would need to standardize the measure units to be the same instead of having one with grams and one with cups and so on and so forth. 


### Generating Tables

Another useful technique for exploratory data analysis is to generate summary tables. 
You may want to use the `dplyr` package in R (`group_by` or `count` functions), as well as the `groupby` and `count` methods in Pandas. [Python example](https://sparkbyexamples.com/pandas/pandas-groupby-count-examples/), [R example](https://dplyr.tidyverse.org/reference/count.html)

[Part 1] Using R and Python, generate a table that shows what **proportion** of recipes contain each type of ingredient, for the most common 20 ingredients.

```{r}
library(dplyr)

# I find the total number of unique recipes
total_recipes <- n_distinct(choc_chip_cookie_ingredients$Recipe_Index)

# Then I generate the top 20 proportions
top_20_ingredients_r <- choc_chip_cookie_ingredients %>%
  group_by(Ingredient) %>%
  summarize(recipe_count = n_distinct(Recipe_Index)) %>%
  mutate(proportion = recipe_count / total_recipes) %>%
  arrange(desc(proportion)) %>%
  slice_head(n = 20)

print(top_20_ingredients_r)
```

```{python}
# First I gotta get the total number of unique recipes
total_recipes = cookies['Recipe_Index'].nunique()

# I calculate the proportions
ingredient_proportions = (
    cookies.groupby('Ingredient')['Recipe_Index']
    .nunique()
    .reset_index(name='recipe_count')
)

ingredient_proportions['proportion'] = ingredient_proportions['recipe_count'] / total_recipes

# Last I get the top 20
top_20_ingredients_py = ingredient_proportions.sort_values('proportion', ascending=False).head(20)

print(top_20_ingredients_py)
```

[Part 2] Print out a character string that lists all of the ingredients that do not appear in at least 20 recipes.

```{r}
rare_ingredients_r <- choc_chip_cookie_ingredients %>%
  group_by(Ingredient) %>%
  summarize(count = n_distinct(Recipe_Index)) %>%
  filter(count < 20) %>%
  pull(Ingredient)

# I collapsed it into a single character string so I could just see them all at once
rare_string_r <- paste(rare_ingredients_r, collapse = ", ")
cat(rare_string_r)
```

```{python}
# I'm going to filter the ingredients appearing in fewer than 20 recipes
rare_ingredients_py = ingredient_proportions[ingredient_proportions['recipe_count'] < 20]['Ingredient'].tolist()

# Again, putting into one string to make it easier to read. Sure enough, the strings are identical in both R and python
rare_string_py = ", ".join(rare_ingredients_py)
print(rare_string_py)
```

### Visualization

Using whatever plotting system you are comfortable with in R or python, see if you can create a couple of useful exploratory data visualizations which address one of the questions you wrote above - or another question which you've come up with as you've worked on this assignment.

[Part 1] Create at least one plot (it doesn't have to be pretty) that showcases an interesting facet of the data.

```{r}
library(ggplot2)
library(dplyr)

# I first need to prepare the data: I want it to count ingredients per recipe and get the rating
recipe_complexity <- choc_chip_cookie_ingredients %>%
  filter(!is.na(Rating)) %>% # Remove the 1,010 missing ratings
  group_by(Recipe_Index) %>%
  summarize(
    num_ingredients = n(),
    Rating = first(Rating)
  )

# 2. Creating the plot in r. 
ggplot(recipe_complexity, aes(x = num_ingredients, y = Rating)) +
  geom_point(alpha = 0.5, color = "steelblue") +
  geom_smooth(method = "lm", color = "red", se = FALSE) + # I added a trend line because I wanted to see it. I made it red for the huskers. 
  labs(
    title = "More Ingredients = Better Quality Cookies?",
    subtitle = "Number of Ingredients vs. Recipe Rating",
    x = "Number of Ingredients",
    y = "Rating (0 to 1)"
  ) +
  theme_minimal()
```

[Part 2] Write 2-3 sentences about what you can learn from that plot and what directions you might want to investigate from here.

Looking at the rating, which is simply from 0-1 though it can be converted to 1-10, there's a very little correlation between the number of ingredients and the overall rating. However, you can see that there are many "outliers" since there is a cluster of 10 ingredient recipes that consistently score very high but there are some recipes that do not. The more ingredients does not determine a rating so I'm curious what causes a recipe with the same amount of ingredients to be "bad". Perhaps it's the type of ingredients, which is something I can investigate further. With that, I could also look into the those "Rare ingredient" recipes to see if they score higher than those with more common ingredients as well. 
